---
title: "GI_F_SanityChecks_Clean_yet"
author: "yet"
date: "2024-11-18"
output: html_document
---

  Description:
  
  MZ: this updates similarly named AR file
We use the sanity checks to make sure GI results are fine and comparable with previous years.

# 0. Preliminary Stuff

```{r Set Environment}
## If you use the drive file format, you can use the code below. If not, put your path here.
options(scipen=999)
Sys.setenv(LANG = "en")
setwd("~/Desktop/BARI/gi_2023")
```

```{r Load Libraries}
library(tidyverse)
library(sf)
library(data.table)
library(dplyr)
```

# 1. Checks
##  checks on TLID files
```{r}
# file with new TLIDs from script 02_PropToTiger_* for 2021
MtchedTLID_2024 <- read.csv("~/Desktop/Academic/BARI/Scripts/GI/Outputs/parcels_updatedTLID_2024.csv")
#98857, 101 in Feb, 2023 #98930, 97 in 2022
sum(is.na(MtchedTLID_2024$TLID)) #124 in 2024 #96 in Sep 23 #1484 in March, 2023
#43 in 2018 #66 in 2019 #65 in 2020 # 1716 in 2021 
## in next chunk I address 2021 missingness 
MtchedTLID_2023 <- read.csv("~/Desktop/Academic/BARI/GI-2023/Outputs_Feb24/parcels_updatedTLID_2023.csv")
```

## TLIDs from previous year
```{r}
# for missing 2021 TLID, replaces with 2020 TLID if known
# We actually want this error 



for(pdl in MtchedTLID_2024[is.na(MtchedTLID_2024$TLID) & !is.na(MtchedTLID_2024$GIS_ID),]$GIS_ID) {
  print(pdl)
  MtchedTLID_2024[MtchedTLID_2024$GIS_ID == pdl & !is.na(MtchedTLID_2024$GIS_ID),]$TLID = MtchedTLID_2023[MtchedTLID_2023$MAP_PAR_ID == pdl & !is.na(MtchedTLID_2023$MAP_PAR_ID),]$TLID
}

# Error in `$<-.data.frame`(`*tmp*`, "TLID", value = integer(0)) : 
# replacement has 0 rows, data has 1

# # for missing 2021 TLID, replaces with 2020 TLID if known
# for(pdl in MtchedTLID_2024[is.na(MtchedTLID_2024[,TLID]) & !is.na(MtchedTLID_2024[,MAP_PAR_ID]),]$MAP_PAR_ID) {
#   print(pdl)
#   MtchedTLID_2024[MtchedTLID_2024$MAP_PAR_ID == pdl & !is.na(MtchedTLID_2024[,MAP_PAR_ID]),]$TLID = MtchedTLID_2023[MtchedTLID_2023[,MAP_PAR_ID] == pdl & !is.na(MtchedTLID_2023[,MAP_PAR_ID]),]$TLID
# }
#From Sep 23
#[1] 2001936000
#[1] 503214000
#[1] 200671000
#[1] 200710000
#[1] 200733000
#[1] 901323500
#[1] 301925000
#[1] 700539001
#[1] 903737150


sum(is.na(MtchedTLID_2024$TLID)) #0 in 2024 #88 in Sep 23 #43 in March, 2023
# 65 in 2020 # 47 in 2021

# save final output
fwrite(data.table(MtchedTLID_2024),"~/Desktop/Academic/BARI/Scripts/GI/Outputs/parcels_updatedTLID_postsanitycheck_2024.csv")
#

```


## check difference
```{r}
setdiff(MtchedTLID_2024$TLID, MtchedTLID_2023$TLID) #152 in 2024 #62 different in Sep 23
#setdiff(MtchedTLID_2021$TLID, MtchedTLID_2020$TLID) #28 different
#setdiff(MtchedTLID_2020$TLID, MtchedTLID_2019$TLID) #64 different
#setdiff(MtchedTLID_2020$TLID, MtchedTLID_2018$TLID) #79 different
#setdiff(MtchedTLID_2019$TLID, MtchedTLID_2018$TLID) #15 different
#sum(MtchedTLID_2020$TLID == 0) #0
#sum(MtchedTLID_2019$TLID == 0) #0
```


## check Parcels files
```{r}
parcels_2024 = fread("~/Desktop/Academic/BARI/Scripts/GI/Outputs/Parcel_final_2024_YET.csv") #98150 #97906 in Sep 23

parcels_2023 = fread("~/Desktop/Academic/BARI/GI-2023/drive/Parcel_final_10032023_YET_postsanitychecks.csv") #98150 

#parcels_2019 <- read.csv("C:/Users/bariuser2/Google Drive/BARI Research #Team Data Library/Geographical Infrastructure/Boston Geographical #Infrastructure 2019/Outputs/Parcel_final_2019_13112019.csv") #98179

#parcels_2018 <- read.csv("C:/Users/bariuser2/Google Drive/BARI Research #Team Data Library/Geographical Infrastructure/Boston Geographical #Infrastructure 2018/Outputs/Parcel_final_2018_11132019.csv") #98136
```

## check descriptive statistics

```{r}
# column type
class(parcels_2023$YR_BUILT)
class(parcels_2023$YR_REMOD)

nrow(parcels_2024[YR_BUILT > YR_REMOD,]) #33 in 2024 # 31 in Sep 23
# 241 for 2023
#241 parcels have the built in 2022 year higher than remod (vs 200 in 2020)

# counts NAs
sum(is.na(parcels_2024$YR_REMOD)) #51605 in 2024
#52507 in Sep 23
#54386 in Feb 2023
#54384 in 2021 #54898 in 2020 #55598 in 2019
sum(is.na(parcels_2024$YR_BUILT)) #13007 in 2024
#13169 in Sep 23
#13511 in Feb 2023
#13511 in 2021 #13827 in 2020 #13998 in 2019

##
length(unique(parcels_2024$Land_Parcel_ID)) #97768 in 2024
#97894 in Sep 23 #98150 in March 2023 #98150 in 2022, #98146 in 2021, #98445 in 2020
#98179 in 2019
length(unique(parcels_2024$TLID)) #13874 in 2024
#13866 in Sep 23 #13826 in March 2023, #13826 in 2022 #13817 in 2021 #13833 in 2020
#13885 in 2019

```

## Comment: Alina (re 2020 files)
: I think we need to add some NSA_Name if existed in 2019

## check properties files
```{r}

#this would be the resulting PAD 2021 from script 03
properties_2024 = fread("~/Desktop/Academic/BARI/Scripts/GI/Outputs/properties_geo_2024_YET.csv")

properties_2023 = fread("~/Desktop/Academic/BARI/GI-2023/drive/properties_geo_10032023_YET_postsanitychecks.csv")
#this would be the resulting PAD 2020 from script 03

#properties_2019 <- read.csv("C:/Users/bariuser2/Google Drive/BARI #Research Team Data Library/Geographical Infrastructure/Boston #Geographical Infrastructure #2019/Outputs/properties_geo_2019_11132019.csv") #this would be the #resulting PAD 2020 from script 03

#properties_2018 <- read.csv("C:/Users/bariuser2/Google Drive/BARI #Research Team Data Library/Geographical Infrastructure/Boston #Geographical Infrastructure #2018/Outputs/properties_geo_2018_11132019.csv") #this would be the resulting PAD 2020 from script 03
```

## Comment: Pad zipcode with zero
## Comment: I think we need to add some NSA_Name if existed in 2019
```{r}
class(properties_2024$YR_REMOD) #integer
class(properties_2024$YR_BUILT) #integer



properties_2024$ZIP_CODE <- as.integer(properties_2024$ZIP_CODE)
class(properties_2024$ZIP_CODE) #integer
class(properties_2023$ZIPCODE) #integer

# pad zipcode with zero
properties_2024$ZIP_CODE = paste("0", properties_2024$ZIP_CODE, sep = "")
table(nchar(properties_2024$ZIP_CODE)) # 5, 177091 # 3 180624 in Sep 23 #3 182239 in 2024

# need shared type for next check
properties_2023$PID = paste(properties_2023$PID)
properties_2024$PID = paste(properties_2024$PID)
View(setdiff(properties_2024$PID, properties_2023$PID)) 
# 2021: 2403 2020: 3091 #4219 in Sep 23 #2054 in 2024

sum(is.na(properties_2024$PID)) #0
sum(is.na(properties_2023$PID)) #0

sum(is.na(properties_2024$GIS_ID)) #0 in Sep 23 #3 in 2023, #3 in 2020
## these 3 properties seem off -- have a south boston nsa,
### share land parcel id with a south boston/seaport building, but
### have a different address (Tremont St) that seems to be South End
sum(is.na(properties_2023$GIS_ID)) #3 in 2023 #0 in 2020

# missing NSA?
sum(is.na(properties_2024$NSA_Name )) #43855 in 2024 #43188 in Sep 2023# 42075 in 2023 # 0: 2021 # previous people did zero which does not make sense
#, I did it NA
#what is NSA name, probably outdated? Toshi comment on Feb 22 2023

table(paste(properties_2024$NSA_Name) %in% properties_2023$NSA_Name[match(paste(properties_2023$Land_Parcel_ID), paste(properties_2024$Land_Parcel_ID))]) 
#in 2024
# FALSE   TRUE 
# 43855 138387 
# in Sep 23
# FALSE   TRUE 
# 43188 137439 
# in Feb 2023
#FALSE   #TRUE 
#42075  135016

# 2021: 0 F 2020: 2331 FALSE
```

## Final Checks 
These checks are compiled from the file "Sanity Checking code for parcels and properties"
```{r}
properties_final <- properties_2024
parcels_final <- parcels_2024

names(properties_final)
n_distinct(properties_final$PID) #182235 in 2024 #180448 in Sep 23 # 176915 in Feb 2023 #174681 in 2021
sum(duplicated(properties_final)) #6 in 2024 #34 in Sep 23 # 3 in 2023
#3 in 2021 #290 in 2020 #216 in 2019 #216 in 2018
sum(duplicated(properties_final$PID)) #7 in 2024 #179 in Sep 23 #176 in Feb 2023 #209 in 2021
# sum(duplicated(cbind(properties_final$PID,properties_final$X,properties_final$Y,properties_final$GIS_ID,properties_final$Land_Parcel_ID,properties_final$ST_NUM,properties_final$ST_NAME,properties_final$UNIT_NUM,properties_final$LU,properties_final$OWN_OCC,properties_final$YR_BUILT,properties_final$YR_REMOD,properties_final$LAND_SF,properties_final$GROSS_AREA,properties_final$NUM_FLOORS,properties_final$unit_N))) #34 in Sep 23 #3 in 2023 #3 in 2021 #290 in 2020
nrow(properties_final) - nrow(distinct(properties_final, PID, X, Y, GIS_ID, Land_Parcel_ID, ST_NUM, ST_NAME, UNIT_NUM, LU, OWN_OCC, YR_BUILT, YR_REMOD, LAND_SF, GROSS_AREA, NUM_FLOORS, unit_N)) #6 in 2024

table(properties_final$LU[duplicated(cbind(properties_final$PID,properties_final$X,properties_final$Y,properties_final$GIS_ID,properties_final$Land_Parcel_ID,properties_final$ST_NUM,properties_final$ST_NAME,properties_final$UNIT_NUM,properties_final$LU,properties_final$OWN_OCC,properties_final$YR_BUILT,properties_final$YR_REMOD,properties_final$LAND_SF,properties_final$GROSS_AREA,properties_final$NUM_FLOORS,properties_final$unit_N))]) # here the CM has 290
#In Sep 23: CC CD CM  E R3 
#            1  3  1  1  1 
#In Sep 23: CD CM  E R3 
#.           1  1  1 31 
# View duplicates
# View(properties_final[duplicated(cbind(properties_final$PID,properties_final$X,properties_final$Y,properties_final$GIS_ID,properties_final$Land_Parcel_ID,properties_final$ST_NUM,properties_final$ST_NAME,properties_final$UNIT_NUM,properties_final$LU,properties_final$OWN_OCC,properties_final$YR_BUILT,properties_final$YR_REMOD,properties_final$LAND_SF,properties_final$GROSS_AREA,properties_final$NUM_FLOORS,properties_final$unit_N)),])
View(properties_final[duplicated(properties_final[, c("PID", "X", "Y", "GIS_ID", "Land_Parcel_ID", "ST_NUM", "ST_NAME", "UNIT_NUM", "LU", "OWN_OCC", "YR_BUILT", "YR_REMOD", "LAND_SF", "GROSS_AREA", "NUM_FLOORS", "unit_N")]), ])

summary(properties_final$unit_N)
View(properties_final[properties_final$unit_N>1000,]) # 2023: none,# 2021: none
# 2020: GIS ID 1303445000 has 1271 units
nrow(properties_final[properties_final$Land_Parcel_ID==300450000,]) 
#866 in 2020; #865 in 2021 #865 in 2023 #865 in Sep 23 #865 in 2024
n_distinct(properties_final$Blk_ID_20) #5156 in 2024 # 5155 in Sep 23 #5158 in Feb 2023 #5478 in 2021 #5479 in 2020
#View(table(properties_final$Blk_ID_10))
n_distinct(properties_final$BG_ID_20) #582 in 2024 #582 in Sep 23 #582 in Feb 2023 #560 in 2021 # 560 in 2020
#View(table(properties_final$BG_ID_10))
#View(table(properties_final$CT_ID_10))

names(parcels_final)
n_distinct(parcels_final$Land_Parcel_ID) # 97768 in 2024 #97894 in Sep 23 #98150 in Feb 2023 #98146 in 2021 #98445 in 2020
summary(parcels_final$property_N)
View(parcels_final[parcels_final$property_N>500,]) #4 have more than 500 in  2024(same as 2020, 2021, 2023)
summary(parcels_final$unit_N)
summary(parcels_final$unit_N_orig)
n_distinct(parcels_final$Blk_ID_20) #5156 in 2024 #5155 in Sep 23 #5156 in Feb 2023, #5479 in 2021
View(table(parcels_final$Blk_ID_20))
n_distinct(parcels_final$BG_ID_20) #582 in 2024 #582 in Feb 2023, #560 in 2021
View(table(parcels_final$BG_ID_20))
n_distinct(parcels_final$CT_ID_20) #209 in 2024 #209 in Feb 2023, #182 in 2021
View(table(parcels_final$CT_ID_20))
View(table(parcels_final$TLID))

max(parcels_final$Land_Parcel_ID) #2205670000 in 2024,2023,2021 and 2020

```

## Save final 2022 files 
```{r}
# parcel 2024
fwrite(parcels_final, "~/Desktop/Academic/BARI/Scripts/GI/Outputs/Parcel_final_postsanitychecked_2024.csv")

# property 2024
fwrite(properties_final, "~/Desktop/Academic/BARI/Scripts/GI/Outputs/properties_geo_postsanitychecked_2024.csv")

```

## Dan's updates 
This chunk is after Dan reviewed the files. making a few updates.
```{r}
# parcel 2022
#parcels_2024 = fread("Parcel_final_20230331_TY_postsanitychecks.csv")
# reorder variables --> full address should be 2nd, not last
setcolorder(parcels_final,c("Land_Parcel_ID","full_address"))
fwrite(parcels_final, "~/Desktop/BARI/GI-2023/outputs/Sanity Checks/Parcel_final_20230911_YET_postsanitychecks.csv")

# property 2021
#properties_2023 = fread("properties_geo_20230331_TY_postsanitychecks.csv")
# drop 3 pure duplicates
properties_final = unique(properties_final)
fwrite(properties_final, "~/Desktop/BARI/GI-2023/outputs/Sanity Checks/properties_geo_20230911_YET_postsanitychecks.csv")
```







